# Bayesian analysis

```{r}
library(PKPDmisc)
library(data.table)
library(tidyverse)
library(knitr)
library(infuser)
```

```{r}
source("../scripts/thin_by.R")
```

Core differences

* Need priors on parameters
* EM algorithms can more robustly handle full block matrices as well as
random effects on less well-defined parameters.

## Priors

Priors in nonmem may be defined in two ways:

1) By distinguishing the thetas/omegas and their priors in `$PRIOR`

* NTHETA=number of Thetas to be estimated
* NETA=number of Etas to be estimated
    * (and to be described by NETAxBETA OMEGA matrix)
* NTHP=number of thetas which have a prior
* NETP=number of Omegas with prior

In our case we have 2 Thetas (TVCL, TVV) and 2 omegas (nCl, nV)

```
$PRIOR NWPRI NTHETA=2 NETA=2, NTHP=2, NETP=2
```

Or the *preferred way* of distinguishing in the block definitions

* $THETA - fixed effects
* $OMEGA - random effects
* $SIGMA - residual error
* $THETAP - theta prior
* $THETAPV - variance of theta prior
* $OMEGAP - omega prior
* $OMEGAPD - degrees of freedom to prior omega matrix

** TODO: fill out explanations of each. **

# Bayesian Problems (and solutions)

Some of the problem areas in bayesian analysis techniques are .....

* autocorrelation
* sensitivity to priors
* ...

## Chains

Technique - a 'master' execution file can be used to generate new paramater values
that can be used for multi-chain runs.

* NSAMPLE - sets number of preturbed parameter estimates, in this case 4
* ISAMPLE - if subsequently estimating, will use the parameters generated from this sample
    * by setting to 0 will not tweak these parameters

```
$EST METHOD=CHAIN FILE=run003chains.chn NSAMPLE=4 ISAMPLE=0 DF=10
```

The larger the DF (degrees of freedom for randomly creating the OMEGAS) the smaller the
difference between newly generated values. Eg, with `DF=10000` almost no difference in values,
whereas DF=5 will have 2x+ differences in values

#### example output comparing output from all chains

** TODO: add more instructions on what all is going on **

```{r}
set.seed(1234567)
modt <- read_file("../modeling/run006c.modt")
map(1:4, ~ write_file(
    infuse(modt, 
           chain_number = ., 
           seed = round(runif(1, 1000, 100000), 0)), 
    file.path("..", "modeling", paste0("run006c", ., ".mod"))))
```

```{r}
r6c1 <- fread("../modeling/run006c1.ext", skip = 1) %>% mutate(chain = 1)
r6c2 <- fread("../modeling/run006c2.ext", skip = 1) %>% mutate(chain = 2)
r6c3 <- fread("../modeling/run006c3.ext", skip = 1) %>% mutate(chain = 3)
r6c4 <- fread("../modeling/run006c4.ext", skip = 1) %>% mutate(chain = 4)
```


```{r}
r6_chains <- bind_rows(r6c1, r6c2, r6c3) %>% 
    filter(ITERATION > 0) %>%
    rename(CL = THETA1,
           V = THETA2,
           EPS = `SIGMA(1,1)`,
           nCL = `OMEGA(1,1)`,
           nV = `OMEGA(2,2)`,
           nCL_nV = `OMEGA(2,1)`)
```

```{r}
kable(head(r6_chains))
```

```{r}
r6_chains %>% select(ITERATION:nV, chain) %>%
    thin_by(10) %>%
    gather(param, value, CL:nV) %>%
    ggplot(aes(x = ITERATION, y = value, color = factor(chain)), alpha = 0.6) +
    geom_line() + facet_wrap(~param, ncol = 2, scales = "free") +
    theme_bw() +
    base_theme() +
    scale_color_discrete(name = "Chain") + 
    theme(legend.position = "bottom") 
```

```{r cache = FALSE}
session_details <- devtools::session_info()
session_details$platform
knitr::kable(session_details$packages)
```
