# Dataset Creation

```{r message=FALSE}
library(knitr)
library(PKPDmisc)
library(tidyverse)
library(mrgsolve) 
```

```{r}
source("../scripts/model_details.R")
```

```{r}
models <- source("../models/models.R")$value
```


## Generate data for mrgsolve

start with a baseline of having 50 individuals worth of data, can scale to 
different amounts of individuals later

```{r}
NIDS <- 50
```

```{r}
demogs <-  data_frame(ID = 1:NIDS)
```

```{r}
for_dosing <- demogs %>% 
                  mutate(
                      CMT = 1, 
                      EVID = 1,
                      TIME = 0,
                      AMT = 1000,
                      RATE = 1000
                  ) 
```

```{r}
one_cmt_iv <- models$use("one_cmt_iv")
```


## Model Details

```{r}
mrgsolve::see(one_cmt_iv)
```

```{r }
one_cmt_iv %>% 
    model_details %>% 
    filter(block != "CAPTURE") %>%
    kable()
```


```{r }
simulated_data <- one_cmt_iv %>% 
    data_set(for_dosing) %>%
    mrgsim(end = 24, delta = 0.25) %>% as_data_frame
```

## Distribution of peak and trough values

* 'peak' defined as 1 hr post infusion and trough 1 hour prior to when next dose would begin
```{r}
simulated_data %>% 
    filter(TIME %in% c(2, 23)) %>% 
    mutate(DV = ifelse(DV < 0, 0, DV)) %>%
    select(ID, DV) %>%
    group_by(ID) %>%
    summarize_all(funs(min, max)) %>%
    gather(sample, value, -ID) %>%
    ggplot(aes(x = value)) + 
    geom_density() +
    facet_wrap(~sample, scales = "free") +
    theme_bw() +
    base_theme()
```

## Predicted Profiles

```{r}
simulated_data %>% 
    filter(IPRED > 0.1) %>%
    ggplot(aes(x = TIME, y = IPRED, group = ID)) +
    geom_line(size = 1.05, alpha = 0.8) + theme_bw() +
    base_theme() + scale_y_log10()
```

## Real world sampling and LLOQ

```{r}

sample_times <- c(2, 4, 8, 16, 24)
LLOQ <- 0.1
sampled_data <- simulated_data %>% 
    filter(TIME %in% sample_times, DV > LLOQ) 

```

Show which, if any, timepoints have bql values and determine percent bql

```{r}
sampled_data %>%
    count(TIME) %>% 
    mutate(baseline = first(n),
           perc_bql = 100 - n/baseline*100) %>%
    filter(perc_bql > 0) %>% 
    select(TIME, perc_bql) %>% knitr::kable()
```

## IPRED and DV vs TIME for all individuals at sampled times

```{r}
list_plots <- sampled_data %>%
    mutate(PNUM = ids_per_plot(ID)) %>%
    split(.$PNUM) %>%
    map(~ 
    ggplot(., aes(x = TIME, y = IPRED, group = ID)) +
    geom_point(aes(y = DV), color = "blue") + 
    geom_line(size = 1.05, alpha = 0.8) + theme_bw() +
    base_theme() + facet_wrap(~ID) +
        scale_y_log10()
    )

print_plots(list_plots)
```

## Prepare for nonmem

```{r}
nm_dat <- sampled_data %>% select(ID, TIME, DV) %>%
    mutate(CMT = 1,
           EVID = 0
           ) %>%
    bind_rows(for_dosing) %>%
        arrange(ID, TIME, desc(EVID))
```

```{r}
kable(head(nm_dat))
```

```{r}
write_nonmem(nm_dat, "../modeling/mdata/simple_nocovar_50id_6tp.csv")
```


```{r cache = FALSE}
session_details <- devtools::session_info()
session_details$platform
knitr::kable(session_details$packages)
```
