# Dataset Creation

```{r message=FALSE}
library(knitr)
library(PKPDmisc)
library(tidyverse)
library(mrgsolve) 
library(infuser)
```

```{r}
source("../scripts/model_details.R")
```

```{r}
models <- source("../models/models.R")$value
```


## Generate data for mrgsolve

start with a baseline of having 50 individuals worth of data, can scale to 
different amounts of individuals later

```{r}
NIDS <- 56
```

```{r}
demogs <-  data_frame(ID = 1:NIDS)
```

```{r}
for_dosing <- demogs %>% 
                  mutate(
                      CMT = 1, 
                      EVID = 1,
                      TIME = 0,
                      AMT = 1000,
                      RATE = 1000,
                      ADDL = 1,
                      II = 12,
                      OBSNUM = 0
                  ) 
```

```{r}
one_cmt_iv <- models$use("one_cmt_iv")
```


## Model Details

```{r}
mrgsolve::see(one_cmt_iv)
```

```{r }
one_cmt_iv %>% 
    model_details %>% 
    filter(block != "CAPTURE") %>%
    kable()
```


```{r }
simulated_data <- one_cmt_iv %>% 
    data_set(for_dosing) %>%
    mrgsim(end = 36, delta = 0.25) %>% as_data_frame
```

## Distribution of peak and trough values

* 'peak' defined as 1 hr post infusion and trough 1 hour prior to when next dose would begin
```{r}
simulated_data %>% 
    filter(TIME %in% c(2, 11)) %>% 
    mutate(DV = ifelse(DV < 0, 0, DV)) %>%
    select(ID, DV) %>%
    group_by(ID) %>%
    summarize_all(funs(min, max)) %>%
    gather(sample, value, -ID) %>%
    ggplot(aes(x = value)) + 
    geom_density() +
    facet_wrap(~sample, scales = "free") +
    theme_bw() +
    base_theme()
```

## Predicted Profiles

```{r}
simulated_data %>% 
    filter(IPRED > 0.1) %>%
    ggplot(aes(x = TIME, y = IPRED, group = ID)) +
    geom_line(size = 1.05, alpha = 0.8) + theme_bw() +
    base_theme() + scale_y_log10(breaks = c(0.1, 1, 10, 20, 30, 40))
```

## Real world sampling and LLOQ

```{r}

sample_times_rich <- c(2, 6, 11, 14, 18, 23)

LLOQ <- 0.1


sampled_data_rich2d <- simulated_data %>% 
    filter(TIME %in% sample_times_rich, DV > LLOQ) %>%
    group_by(ID) %>% mutate(OBSNUM = dplyr::row_number(ID))

sparser_scenarios <- list(
    "s2d" = c(1, 3, 4, 6),
    "s1dpt" = c(1, 3, 6),
    "r1d" = c(1, 2, 3),
    "s2trough" = c(3, 6),
    "s1d" = c(1, 3),
    "tr" = c(3)
)

sparser_scenario_data <- map(names(sparser_scenarios), function(scenario) {
       res <- sampled_data_rich2d %>%
         filter(OBSNUM %in% sparser_scenarios[[scenario]])
       return(setNames(res, scenario))
})
```


## IPRED and DV vs TIME for all individuals at sampled times

```{r}
list_plots_rich <-  simulated_data %>%
    select(-DV) %>% 
    left_join(sampled_data_rich2d %>% select(ID, TIME, DV)) %>%
    mutate(PNUM = ids_per_plot(ID)) %>%
    split(.$PNUM) %>%
    map(~ 
    ggplot(., aes(x = TIME, y = IPRED, group = ID)) +
    geom_point(aes(y = DV), color = "blue") + 
    geom_line(size = 1.05, alpha = 0.8) + theme_bw() +
    base_theme() + facet_wrap(~ID) +
        scale_y_log10(breaks = c(0.1, 1, 5, 10, 20, 30, 40))
    )

print_plots(list_plots_rich)
```

## Prepare for nonmem

```{r}
nm_dat_rich <- sampled_data_rich2d %>% select(ID, TIME, DV, OBSNUM) %>%
    mutate(
        CMT = 1,
        EVID = 0
    ) %>%
    bind_rows(for_dosing) %>%
        arrange(ID, TIME, desc(EVID))
```

*rich data

```{r}
kable(head(nm_dat_rich, n = 14))
```

## Create chains

```{r}
modt <- read_file("../modeling/run007c.modt")
BASE_MODEL_NUM <- 7
scenario_df <- as_data_frame(
        expand.grid(chain = 1:4, scenario = names(sparser_scenarios))
    ) %>% arrange(scenario) %>%
    mutate(scenario_num = BASE_MODEL_NUM + as.numeric(as.factor(scenario)))

kable(scenario_df)
```

Want to inject in the chain number and scenario name, as well as subset
the data relevant to the scenario. The ignore statements control which 
OBSNUM will be retained for estimation

```{r}
by_row(scenario_df, function(row) {
    set.seed(1234567)
    ignore_obs <- setdiff(1:length(sample_times_rich), sparser_scenarios[[row$scenario]])
    ignore <- paste0("IGNORE=(OBSNUM.EQN.", ignore_obs, ")", collapse = " ")
    write_file(
    infuse(modt, 
           chain_number = row$chain, 
           seed = round(runif(1, 1000, 100000), 0),
           ignore = ignore,
           scenario = row$scenario), 
    file.path("..", "modeling", paste0("run", pad_left(row$scenario_num, 3), "c", row$chain, ".mod")))
    })

```

## output data

```{r}
write_nonmem(nm_dat_rich, "../modeling/mdata/simple_nocovar_56id_6tp_md.csv")
```


```{r cache = FALSE}
session_details <- devtools::session_info()
session_details$platform
knitr::kable(session_details$packages)
```
