# impact of dispersion (DF) on individual parameter estimates

```{r}
library(tidyverse)
library(infuser)
library(PKPDmisc)
library(future)
```

```{r }
read_nm <- function(.file) {
    data.table::fread(.file, skip = 1)
}

# extract the lower matrix into a list of rows
extract_lower <- function(.omega) {
    if (length(.omega) == 1) {
        return(list(.omega))
    }
    lower_tri <- lower.tri(.omega, diag = T)
    lapply(1:nrow(.omega), function(i){
        .omega[i, ][lower_tri[i, ]]
    })
}

# collapse a matrix to a string representing the lower triangular matrix, 
# optionally with fix on the first element, which is required for priors 
collapse_to_lower <- function(.omega, fix = TRUE) {
    lower_list <- extract_lower(.omega)
    if (fix) {
        lower_list[[1]] <- paste(lower_list[[1]], "FIX")
    }
    paste(lapply(lower_list, 
                 paste, collapse = " "), 
          collapse = "\n ")
}

#' @param run run name
#' @param pv_sf prior variance scaling factor - defaults to 1 but can increase or decrease prior variance on theta
process_run <- function(run, pv_sf = 1) {
    ex_ext <- read_nm(paste0(run, ".ext")) %>% filter(ITERATION == -1000000000)
    
    # get the parameter values to use as initial estimates for each structure,
    # and convert the parameter sets that are lower triangular matrices to 
    # that format
    tv_theta <- ex_ext %>% select(contains("THETA"))
    tv_omega <- mrgsolve::as_bmat(ex_ext %>% select(contains("OMEGA")))[[1]] 
    tv_sigma <- mrgsolve::as_bmat(ex_ext %>% select(contains("SIGMA")))[[1]]
    
    cov_mat <- read_nm(paste0(run, ".cov")) %>% select(-NAME) 
    nm_cov_mat <- names(cov_mat)
    cov_mat <- as.matrix(cov_mat)
    theta_indices <- which(grepl("THETA", nm_cov_mat))
    
    # get the prior variance for theta  
    theta_pv <- collapse_to_lower(cov_mat[theta_indices, theta_indices]*pv_sf)
    # the priors for omega and sigma are the same as the tv's, however
    # must also add FIX
    return(list(
        theta = paste("(0.001, ", tv_theta, ")", collapse = "\n"),
        omega = collapse_to_lower(tv_omega, fix = FALSE),
        sigma = collapse_to_lower(tv_sigma, fix = FALSE),
        thetap = paste(tv_theta, "FIX", collapse = "\n"),
        thetapv = theta_pv,
        omegap = collapse_to_lower(tv_omega),
        sigmap = paste(tv_sigma, "FIX", collapse = "\n")
    ))
}
```


## Influence of dispersion factor (DF) on sigma and omega, for purposes of estimation

Parameter values are taken from the 36 and 56 individual fits, in which the estimated
values are similar, however, the prior variance on theta and omega is such that
the 56 individual group has approximately 2x the variance. 

In combination with tweaking the dispersion factors, this should give some indication
of the relative contributions of dispersion factor vs prior variance, leading to 
more targeted efforts on which parameters to tune to optimize the behavior for
the priors to balance incoming data for an individual, against prior knowledge.


```{r }
run113 <- process_run("../modeling/rundir_001/run113_nids_36_s2d_c1")
run119 <- process_run("../modeling/rundir_001/run119_nids_56_s2d_c1")
```


## Parameters to replace

```{r }
template_mod <- read_file("../modeling/rundir_002/run000.modt")
infuser::variables_requested(template_mod)
```


```{r }

sparser_scenarios <- list(
    "s2d" = c(1, 3, 4, 6),
    "s1dpt" = c(1, 3, 6),
    "r1d" = c(1, 2, 3),
    "s2trough" = c(3, 6),
    "s1d" = c(1, 3)
)

sample_times_rich <- c(2, 6, 11, 14, 18, 23)

```


```{r }
START_RUNNUM <- 0 # will be added to index, so this will start at run001
all_scenario_info <- expand.grid(SIGMAPD = c(20, 50, 100, 500, 1000, 10000),
                                  OMEGAPD = c(3, 20, 50, 100, 500),
                                 sampling = names(sparser_scenarios),
                                  data = paste0("ind_", pad_left(1:56, 2), ".csv")
                                 ) %>%
    as_tibble %>% 
    mutate(scenario = paste0("ODF_", OMEGAPD, "_SDF_", SIGMAPD, "_", sampling),
           seed = round(runif(n(), 10000, 100000), 0)) %>%
    arrange(data, scenario) %>%
    group_by(data) %>%
    mutate(runnum = 1:n())

res <- all_scenario_info %>% ungroup %>% slice(3L) %>%
    by_row(function(scenario_info) {
        set.seed(1234567)
        ignore_obs <- setdiff(1:length(sample_times_rich), sparser_scenarios[[scenario_info$sampling]])
        ignore <- paste0("IGNORE=(OBSNUM.EQN.", ignore_obs, ")", collapse = " ")
        scenario_info$ignore <- ignore
        ind_dir <- file.path("..", 
                  "modeling", 
                  "rundir_002", 
                  tools::file_path_sans_ext(scenario_info$data))
        if (!dir.exists(ind_dir)) {
            dir.create(ind_dir)
        }
        runname <- paste0("run", pad_left(scenario_info$runnum, 3), "_", scenario_info$scenario)
        write_lines(infuse(template_mod, c(PKPDmisc::capitalize_names(run113), scenario_info)),
                              file.path(ind_dir, paste0(runname, ".mod"))
                              )
    }, .collate = "list", .labels = F) 
```

## individual data

```{r}
all_data <- read_nonmem("../modeling/mdata/simple_nocovar_56id_6tp_md.csv")

res <- all_data %>% split(.$ID) %>%
    map(~ write_nonmem(., file.path("..", "modeling", "mdata", paste0("ind_", pad_left(unique(.$ID), 2), ".csv")))) 
```

